---
title: "01_data_analysis"
output: html_document
date: "2022-11-21"
editor_options: NR
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Analysing fishing effort and demand data to find a relationship between the two variables. 

Fishing effort data from Camilla - Global data-set built for regional models 
Demand data from FAO Food Balance Sheet downloaded on 21/11/2022

## load data 


```{r load libraries and data}

rm(list=ls())
library(tidyverse)

summarise <-dplyr::summarise

effort <- read.csv(file = "effort_histsoc_1841_2010_regional_models.csv") 
catch <- read.csv(file = "calibration_catch_histsoc_1850_2004_regional_models.csv")
fao<-read.csv("Human forcings/FAOSTAT_data_en_10-27-2022.csv")

```

## subsetting to needs 

```{r data exploration}

# subset SE au and the demersal group 
head(effort)
unique(effort$FGroup) # see Sea Around Us website for FGroup codes and meaning 
range(effort$Year)

effort_au<-effort %>% 
  filter(region == "SEAustralia.mizer", Sector == "Industrial", FGroup %in% c("flatfish<90cm", "demersal<30cm","demersal>=90cm", "flatfish>=90cm","demersal30-90cm")) %>% 
  filter(Year >= 1961) %>% 
  group_by(Year) %>% 
  summarise(effort = sum(NomActive)) %>% 
  ungroup()


# subset SE au and the demersal group for FAO data 

head(fao)
unique(fao$Domain.Code) # only fish
unique(fao$Element) # interested in production only as a proxy for consummption which is  aproxy for demand - WARNING CHECK ! production or supply?? 
unique(fao$Item)
unique(fao$Flag.Description)

range(fao$Year)

fao_au<-fao %>% 
  filter(Year <= 2010) %>% 
  filter(Element %in% c("Production"), Item %in% c("Demersal Fish")) %>% 
  dplyr::select(Year, Value) # Item, Element, Value, Unit)

head(fao_au)
```



## percentage changes 

```{r perc}

ref<-effort_au %>% 
  filter(Year >=1961, Year <=1970) %>% 
  # filter(Year >=1990, Year <=1999) %>% 
  summarise(ref = mean(effort))

# 1998 414% compared to 1960s # value in mean 60s = 3762976, value in 1998= 19358216
# 1998 31% compared to 1990s # value in mid 90s = 14720592; value in 1998= 19358216
effort_au<-effort_au %>% 
  mutate(perc = ((effort - ref$ref)/ref$ref)*100)

View(effort_au)

ref<-fao_au %>% 
  # filter(Year >=1961, Year <=1970)
  filter(Year >=1990, Year <=1999)%>% 
  summarise(ref = mean(Value))

fao_au<-fao_au %>% 
  mutate(perc = ((Value - ref$ref)/ref$ref)*100)

```

## plot data 

```{r plot}

my_theme<-theme_bw()+
  theme(text = element_text(size = 10), # this should be overwritten by the below
        plot.title = element_text(size = 10),
        axis.title = element_text(size = 9),
        axis.text = element_text(size = 6),
        legend.title=element_text(size = 9),
        legend.text=element_text(size = 8),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.key.size = unit(0.1, "cm"))
 
plot_effort<-ggplot(effort_au, aes(x=Year, y=effort)) +
  # geom_hline(yintercept = 0, linetype = "dashed", color = "grey50", size = 0.3)+
  geom_point(size = 1) +
  geom_line()+
  # scale_color_manual(values = c("red", "blue"))+
  # scale_shape_manual(values = c(16,15))+
  # coord_flip() +
  # ylim(-1,1)+
  my_theme+
  xlab("Year")+
  ylab("Effort (Kw*Day)")+
  ggtitle("Effort")
# facet_grid(MEM~ESM)

plot_effort


plot_fao<-ggplot(fao_au, aes(x=Year, y=Value)) +
  # geom_hline(yintercept = 0, linetype = "dashed", color = "grey50", size = 0.3)+
  geom_point(size = 1) +
  geom_line()+
  # scale_color_manual(values = c("red", "blue"))+
  # scale_shape_manual(values = c(16,15))+
  # coord_flip() +
  # ylim(-1,1)+
  my_theme+
  xlab("Year")+
  ylab("Production (tonness)")+
  ggtitle("Production")
# facet_grid(MEM~ESM)

plot_fao

library("patchwork")

jpeg("Effort_production.jpg", width = 5, height = 6, units = "in", res = 300, bg = "transparent")
plot_effort/plot_fao
dev.off()

# plot relative values 

# combine  data
# 1 create a column id and standardise column names across datasets 

fao_au$Data<- "Production"
effort_au$Data<-"Effort"
all_data<-fao_au %>% 
  full_join(effort_au)

all_data

plot_all<-ggplot(all_data, aes(x=Year, y=perc, color =Data, shape = Data)) +
  # geom_hline(yintercept = 0, linetype = "dashed", color = "grey50", size = 0.3)+
  geom_point(size = 1) +
  geom_line()+
  scale_color_manual(values = c("red", "blue"))+
  scale_shape_manual(values = c(16,15))+
  # coord_flip() +
  # ylim(-1,1)+
  my_theme+
  xlab("Year")+
  ylab("Change (%)")+
  ggtitle("Change in demersal fishing effort and demersal fish production relative to 1990-1999")
# facet_grid(MEM~ESM)

plot_all

jpeg("Effort_production_together1990.jpg", width = 6, height = 5, units = "in", res = 300, bg = "transparent")
plot_all
dev.off()

# WARNING - this plot should be done using the values of effort and production, not the % changes. 
# start from the 2 separate dataset. 
all_data2<-all_data %>% 
  select(Year, perc, Data) %>% 
  spread(key=Data, value =perc)

plot_all_correlation<-ggplot(all_data2, aes(x=Effort, y = Production)) +
  # geom_hline(yintercept = 0, linetype = "dashed", color = "grey50", size = 0.3)+
  geom_point(size = 1) +
  # geom_line()+
  # scale_color_manual(values = c("red", "blue"))+
  # scale_shape_manual(values = c(16,15))+
  # coord_flip() +
  # ylim(-1,1)+
  my_theme+
  xlab("Effort (%)")+
  ylab("Production (%)")+
  ggtitle("Change in demersal fishing effort and demersal fish production relative to 1990-1999")
# facet_grid(MEM~ESM)

plot_all_correlation

```

###Using the original values to plot the correlation 

```{r}


effort_au$Value <- effort_au$effort

fao_au$Data<- "Production"
effort_au$Data<-"Effort"


full_data<-fao_au %>% 
  full_join(effort_au)

full_data2<-full_data %>% 
  select(Year, Value, Data) %>% 
  spread(key=Data, value =Value)

#Plotting all correlation

my_theme<-theme_bw()+
  theme(text = element_text(size = 10), # this should be overwritten by the below
        plot.title = element_text(size = 10),
        axis.title = element_text(size = 9),
        axis.text = element_text(size = 6),
        legend.title=element_text(size = 9),
        legend.text=element_text(size = 8),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.key.size = unit(0.1, "cm"))

plot_full_correlation<-ggplot(full_data2, aes(x=Effort, y = Production)) +
  # geom_hline(yintercept = 0, linetype = "dashed", color = "grey50", size = 0.3)+
  geom_point(size = 1) +
  # geom_line()+
  # scale_color_manual(values = c("red", "blue"))+
  # scale_shape_manual(values = c(16,15))+
  # coord_flip() +
  # ylim(-1,1)+
  my_theme+
  xlab("Effort (Kw*Day)")+
  ylab("Production (tonnes)")+
  ggtitle("Change in demersal fishing effort and demersal fish production relative to 1961-2010")
# facet_grid(MEM~ESM)

plot_full_correlation


#jpeg("Effort_production_correlation1961-2010.jpg", width = 6, height = 5, units = "in", res = 300, bg = "transparent")
#plot_full_correlation
#dev.off()

```

## statistical analysis to quantify relationship. e.g. linear model 

```{r linear model}

#putting the data together data frame
df <- data.frame( x= full_data2$Effort,
       y = full_data2$Production)
			

# fit linear model
linear_model <- lm(y ~ x^2, data=df)

# view summary of linear model
summary(linear_model)

plot(linear_model)

plot_full_correlation
abline(linear_model)

# this is a simple linear model to check the correlation - let us now plot the regression line in ggplot 

```


##PLotting single linear regression line in a ggplot 

```{r single linear regression}



plot_regression<-ggplot(full_data2, aes(x=Effort, y = Production)) +
  # geom_hline(yintercept = 0, linetype = "dashed", color = "grey50", size = 0.3)+
  geom_point(size = 1) +
  # geom_line()+
  # scale_color_manual(values = c("red", "blue"))+
  # scale_shape_manual(values = c(16,15))+
  # coord_flip() +
  # ylim(-1,1)+
  geom_smooth(method='lm', se = FALSE)+
  my_theme+
  xlab("Effort (Kw*Day)")+
  ylab("Production (tonnes)")+
  ggtitle("Linear Regression Plot")
  
  
  plot_regression
  
#jpeg("Linear_Regression_Plot_Effort_and_Production.jpg", width = 6, height = 5, units = "in", res = 300, bg = "transparent")
#plot_regression
#dev.off()


  
  # alternatively use co-eeficient and intercept 
  
  #plot_regression2<-ggplot(full_data2, aes(x=Effort, y = Production)) +
  # geom_hline(yintercept = 0, linetype = "dashed", color = "grey50", size = 0.3)+
  #geom_point(size = 1) +
  # geom_line()+
  # scale_color_manual(values = c("red", "blue"))+
  # scale_shape_manual(values = c(16,15))+
  # coord_flip() +
  # ylim(-1,1)+
 #geom_abline(slope = coef(linear_model) [["Production"]],
             #intercept = coef(linear_model) [["(Intercept)"]]) +
  #my_theme+
  #xlab("Effort (Kw*Day)")+
  #ylab("Production (tonnes)")+
  #ggtitle("Linear Regression PLot") 
  
  #did not work : Error - "Error in coef(linear_model)[["Production"]] : subscript out of bounds"
  


```

##PLotting multiple linear regression line in a ggplot 

```{r multiple regression}

df2 <- full_data2

# create multiple linear model


lm_fit <- lm(Production~ Effort, data=df2)
summary(lm_fit)

# save predictions of the model in the new data frame 
# together with variable you want to plot against
predicted_df <- data.frame(pred = predict(lm_fit, df2), production = df2$Production) 

# this is the predicted line of multiple linear regression

plot_multiple<-ggplot(data = df2, aes(x=Effort, y = Production)) +
  # geom_hline(yintercept = 0, linetype = "dashed", color = "grey50", size = 0.3)+
  geom_point(size = 1) +
  geom_line( data = predicted_df, aes(x=pred, y= production))+
  # scale_color_manual(values = c("red", "blue"))+
  # scale_shape_manual(values = c(16,15))+
  # coord_flip() +
  # ylim(-1,1)+
  my_theme+
  xlab("Effort (Kw*Day)")+
  ylab("Production (tonnes)")+
  ggtitle(" Multiple Linear Regression PLot")

plot_multiple

# getting a single line - doesn't make sense : check the formula again 

```


