---
title: SEA Model 
author: NR
editor_options: 
  chunk_output_type: console
---

Setting up Libraries

```{r}
#Load data 
rm(list=ls())

#Load allll the libraries - use install packages if library doesn't work
library(readxl)
library(tidyverse)
library(devtools)
library(plyr) 
library(Rcpp)
library(reshape2) 
library(inline)
library(ggplot2)
library(ggthemes)
library(vegan)
library(mizer)
library(patchwork)
library(cowplot)
library(ggpubr)
library(matrixStats)
library(mizerMR)
library(mizerExperimental)
library(parallel)
library(tictoc)
library(readxl)
library(mizerHowTo)
```

Loading param from Cami and formatting it to Mizer
```{r}
DFparam <- read_excel("mixed fihseries data/DFparam.xlsx")

Species_params <- DFparam[,(-c(1,10,15,16,17,18,19,20,21,25,26,27,28,29,30,32,34,35,36,37))]
Species_params$R_max <- Species_params$r_max


Species_params<- Species_params[,(-c(15))]
Species_params$biomass_observed <- Species_params$ssbObs
avail_pp<- c(1.0, 0.2, 0.8, 0.2, 0.8, 0.2, 0.0, 0.3, 0.2, 0.2, 0.3, 0.2, 0.2, 0.3, 0.2, 0.2, 0.0, 0.0, 0.0)
avail_bb <- c(0.0, 0.4, 0.0, 0.4, 0.0, 0.4, 0.8, 0.5, 0.4, 0.4, 0.5, 0.4, 0.4, 0.5, 0.4, 0.4, 0.8, 0.8, 0.8)
Species_params$interaction_resource<- avail_pp+avail_bb
Species_params$ssbObs <- NULL

Species_params$l_inf <- exp(log(Species_params$w_inf/Species_params$a)/Species_params$b)

Species_params$FGroup<- c("bathypelagic<30cm", "demersal30-90cm", "demersal30-90cm","bathydemersal30-90cm", "benthopelagic30-90cm", "benthopelagic30-90cm", "shark>=90cm","demersal30-90cm","flatfish<90cm", "benthopelagic30-90cm", "benthopelagic30-90cm", "bathypelagic30-90cm", "benthopelagic30-90cm", "benthopelagic>=90cm", "benthopelagic>=90cm", "bathydemersal>=90cm", "shark>=90cm", "shark>=90cm", "shark>=90cm" )

# Fishing params 
Gear_params <- DFparam[,c(2,17,18,19,20,21,28)]

trawl <- function(w,W25_S,W50_S,...){

  SR <- W50_S - W25_S
  S1 <- W50_S*log(3)/SR
  S2 <- S1 / W50_S
  sel <-1 / (1 + exp(S1 - S2*w))
  return(sel)
}

gillnet<-function(w,W50_B,sigma_B,...){
  sel<-exp(-(w-W50_B)^2/sigma_B^2)
  return(sel)
}

# not working for now ....
unique(Gear_params$species)
demersal <-c("centroberyx affinis",  "nemadactylus macropterus","platycephalus richardsoni", "zeus faber" , "seriolella brama","hoplostethus atlanticus", "macruronus novaezelandiae", "seriolella punctata", "rexea solandri","genypterus blacodes")
pelagic <- c("myctophids","sillago flindersi","nototodarus gouldi","helicolenus barathri","trachurus declivis","squalus spp.", "pristiophorus cirratus","mustelus antarcticus", "galeorhinus galeus" )

str(Gear_params)
Gear_params$gear=ifelse(Gear_params$species %in% demersal, "demersal", "pelagic")
# end not working 
 
Gear_params$gear<-"pelagic"
Gear_params[c(6,8:16),'gear']<-"demersal"
head(Gear_params)
```



```{r}
#Gear_params$sel_func<- "sigmoid_length" 

#Gear_params$l25<- exp(log(Gear_params$W25_S/Species_params$a)/Species_params$b)
#Gear_params$l50<- exp(log(Gear_params$W50_S/Species_params$a)/Species_params$b)



params <- newMultispeciesParams(species_params = Species_params, gear_params = Gear_params, kappa = 0.002191539)


```

```{r}
plotSpectra(params)
Simparam <- project(params, effort = 1)
plotSpectra(Simparam)
plotBiomass(Simparam)


```

Code here shows that the biomass is decreasing drastically - need to change the R_max value
```{r}

params@species_params$R_max<- params@resource_params$kappa*params@species_params$w_inf^-1
Simparam<-project(params, effort = 1)
plotSpectra(Simparam)
plotBiomass(Simparam)

```

All Species Co-exist : Except one species - Squalus spp.

```{r}
plotFeedingLevel(Simparam, include_critical = TRUE)
plotGrowthCurves(Simparam, species_panel = TRUE)
plotlySpectra(Simparam)

plotlyPredMort(Simparam)
plotDiet(Simparam)#, species = "squalus spp.")
plotEnergyBudget(Simparam)
```

Loading and plotting : catch and effort (Fish-MIP data)


```{r}
effort_his <- read.csv(file = "effort_histsoc_1841_2010_regional_models.csv") 
catch_his <- read.csv(file = "calibration_catch_histsoc_1850_2004_regional_models.csv")

effort_SEA <- filter(effort_his,region=="SEAustralia.mizer")#subset - only industrial , all fishing countries, gear - trawl bottom and gillnet (check the  Fgroups), plot per gear and Fgroups- facet_wrap().

effort_tot <- effort_SEA


effort_tot<- effort_tot %>% 
  group_by(Year,FGroup) %>% 
  dplyr::summarise(TotalEffort=sum(NomActive)) %>% 
  ungroup()

 head(effort_tot)

ggplot(data = effort_tot, aes(x=Year,y=TotalEffort, color = FGroup))+
  geom_line()

catch_SEA <- filter(catch_his, region=="SEAustralia.mizer")

catch_tot<- catch_SEA
catch_tot$Totalcatches <- catch_tot$Reported+catch_tot$IUU+catch_tot$Discards
 head(catch_tot)#checked-works
 
catch_tot <- catch_tot %>%
  group_by(Year) %>%
  dplyr::summarise(Totalcatches=sum(Totalcatches))

ggplot(data = catch_tot, aes(x=Year,y=Totalcatches))+
  geom_line()

#ggplot(data = effort_SEA, aes(x= Year, y= NomActive, group = 1))
#+geom_line(linetype="solid", color="forestgreen")+ theme_minimal()+labs(x="Year" ,y="Fishing effort", title = "Fishing effort (1841-2010)- SE-A")+theme(plot.title = element_text(hjust = 0.5,size = 10,face = "bold"))

#ggplot(data = catch_SEA, aes(x= Year, y= Reported, group = 1))+geom_line(linetype="solid", color="turquoise4")+ theme_minimal()+labs(x="Year" ,y="Catch (reported)", title = "Catches(1920-2004)- SE-A")+theme(plot.title = element_text(hjust = 0.5,size = 10,face = "bold"))



```
Creating a subset of effort and catch from the above fishmip data - with only industrial sector and trawl bottom and gillnet gear.

```{r}

# try filter only cephalopods, demersal, bathydemersal, flatfish, rays, shark

effort_sub <- filter(effort_SEA, Sector == "Industrial",Gear =="Trawl_Bottom" | Gear == "Gillnets", FGroup %in% c("cephalopods", "bathydemersal<30cm", "bathydemersal>=90cm","bathydemersal30-90cm", "demersal<30cm", "demersal>=90cm", "demersal30-90cm", "flatfish<90cm", "rays<90cm", "rays>=90cm", "shark>=90cm", "bathypelagic<30cm", "bathypelagic30-90cm", "benthopelagic30-90cm", "benthopelagic>=90cm"))
# Gear %in% c("Trawl_Bottom", "Gillnets")

head(effort_sub)

effort_tt<- effort_sub


effort_tt<- effort_tt %>% 
  group_by(Year,Gear,FGroup) %>% 
  dplyr::summarise(TotalEffort=sum(NomActive)) %>%
  ungroup()

head(effort_tt)

ggplot(data = effort_tt, aes(x = Year, y = TotalEffort,color = FGroup)) +
  geom_line( size = 1) +
  labs(title = "Gear") + 
  facet_wrap(~ Gear)


catch_sub <- filter(catch_SEA,FGroup %in% c("cephalopods", "bathydemersal<30cm", "bathydemersal>=90cm","bathydemersal30-90cm", "demersal<30cm", "demersal>=90cm", "demersal30-90cm", "flatfish<90cm", "rays<90cm", "rays>=90cm", "shark>=90cm", "bathypelagic<30cm", "bathypelagic30-90cm", "benthopelagic30-90cm", "benthopelagic>=90cm"))

head(catch_sub)

catch_tt<- catch_sub
catch_tt$Totalcatches <- catch_tt$Reported+catch_tt$IUU+catch_tt$Discards

catch_tt<- catch_tt %>%
  group_by(Year,FGroup) %>%
  dplyr::summarise(Totalcatches=sum(Totalcatches)) %>%
  ungroup()

head(catch_tt)

ggplot(data = catch_tt, aes(x = Year, y = Totalcatches,color = FGroup)) +
  geom_line( size = 1) +
  labs(title = "FGroup") 
  
#ggplot(data = catch_tt, aes(x = Year, y = Totalcatches,color = FGroup)) + geom_line( size = 1) +labs(title = "SAUP") +facet_wrap(~ SAUP) - each SAUP has only 1 observation - too many plots with too little data .  


  #  platycephalus - flatfish , 17-19,7 - sharks, 3 - ceph, 6- demersal <30, - match species to Fgroup and gear. - set up gear params- gear/year- 18
 

```

The aim of this script is to analyse fishing effort and demand data to find a relationship between the two variables and to develop future scenarios of fishing effort based on demand. 

Data origin: 
1. Fishing effort data from Rousseau et al (in prep), aggregated and extrapolated by Camilla Novaglio for FisHMIP use.  
2. Demand data from FAO Food Balance Sheet downloaded on 21/11/2022
3. Population and GDP data from ISIMIP server, e.g.: 
scp -r b381217@levante.dkrz.de:/work/bb0820/ISIMIP/ISIMIP3b/InputData/socioeconomic/pop/ssp1/population_ssp1_national_annual_2015_2100.csv /Users/camillan/Desktop/FAO_report_plots/chp3_socioeconomy/Data/

Steps: 
1. Consider fishing effort and demand for Australia (or for South East Australia, depending on data aggregation) demersal fish groups.
2. Investigate the relationship between fishing effort an demand using linear and additive models. Is the rate of change in effort proportional to the rate of change in demand? 
3. Consider two scenarios of population and GDP, SSP1-mitigation and SSP5-high-emission, for Australia. 
4. Following the approach in Naylor et al. 2021 (), calculate future demand for demersal fish in Australia based on current demand and future growth in population and GDP in Australia: 
$$
C1 = C0*(Y1/Y0) + eps*((G1 - G0)/G0)
$$
Where C1 = demand in t+1, C0 = current demand, Y0 = current population, Y1 = population in t+1, eps = elasticity, G0 = current GDP, G1 = GDP in t+1. 
5. Building on the approach in Naylor et al. 2021, calculate future fishing effort for demersal fish in Australia based on current effort and future growth in demand. the justification for using this approach is that there is a linear relationship between demand and effort as demonstrated in point 2: 
$$
E1 = E0*(C1/C0)
$$
Where E1 = effort in t+1 and E0 = current effort. 

Outputs: 
1. Plots of demand and effort data and projections.  
2. Data. Use food_final adn effort_final as forcing for regional model of South East Australia. 

## load packages and themes  

```{r load libraries and data}

rm(list=ls())

library(tidyverse)
library(mgcv)
library(scales)
library(patchwork)

summarise <-dplyr::summarise

my_theme<-theme_bw()+
  theme(text = element_text(size = 10), # this should be overwritten by the below
        plot.title = element_text(size = 10),
        axis.title = element_text(size = 9),
        axis.text = element_text(size = 6),
        legend.title=element_text(size = 9),
        legend.text=element_text(size = 8),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.key.size = unit(0.1, "cm"))


dir = "/Users/raveendransankaran/Documents/prepare-Nilar14/"

```

## effort and demand data for Australia  

```{r data exploration}

effort <- read.csv(file = paste0(dir,"effort_histsoc_1841_2010_regional_models.csv"))
fao<-read.csv(file = paste0(dir, "FAOSTAT_data_en_10-27-2022.csv"))

# # subset SE au and the demersal group 

effort_au<-effort %>% 
  filter(region == "SEAustralia.mizer", Sector == "Industrial", FGroup %in% c("flatfish<90cm", "demersal<30cm","demersal>=90cm", "flatfish>=90cm","demersal30-90cm")) %>% 
  filter(Year >= 1961) %>% 
  group_by(Year) %>% 
  summarise(effort = sum(NomActive)) %>% 
  ungroup()

# # subset SE au and the demersal group for FAO data 
# unique(fao$Element) # following Naylor et al 2021, consider "Food supply quantity (kg/capita/yr)" as per-capita consumption/demand.  

fao_au<-fao %>% 
  filter(Year <= 2010) %>% 
  filter(Element %in% c("Food supply quantity (kg/capita/yr)"), Item %in% c("Demersal Fish")) %>%  # OR Food ?
  dplyr::select(Year, Item, Element, Value, Unit)

```

## relationship between effort and demand 

There is a positive relationship between effort and demand. This relationship is stronger when considering the *Food* category compared to the *Food supply quantity (kg/capita/yr)* category. The relationship explains about 50% of the variability in the data. While we use *Food supply quantity (kg/capita/yr)* to calculate future demand based on population and GDP, we multiply *Food supply quantity (kg/capita/yr)* by *population* and obtain *Food* to calculate future effort based on demand. 

```{r}

fao_au<-fao_au %>% 
  mutate(Data = "Food") %>% 
  select(Year, Value, Data)

effort_au<- effort_au %>% 
  mutate(Data = "Effort") %>% 
  rename(Value = effort) %>%
  select(Year, Value, Data)

all_data<-fao_au %>% 
  full_join(effort_au) %>% 
  spread(value = Value, key = Data) 

ggplot(all_data, aes(x = Food, y = Effort))+
  geom_point()+
  geom_smooth(method = "lm")

m<-lm(Food~Effort, data = all_data)
summary(m) 

mb<-gam(Food~Effort, data = all_data)
summary(mb) 

# all data 
all_data <-all_data %>% 
  `colnames<-`(c("year", "effort", "food")) %>%
  mutate(year = as.character(year)) # to match pop and gdp format

```

## popoulation and GDDP data for Australia 

```{r}

# POPULATION
# changed format of files: delete first rows for readability: 
#
# national population in 10^6 people
# (commas in the country name have been replaced by ;)
#

pop_hist <- read.csv(file = paste0(dir,"population_histsoc_national_annual_1950_2014.csv")) %>% 
  filter(country == "Australia") %>% 
  select(-iso, -country) %>% 
  gather(value = pop, key = year) %>% 
  mutate(year = gsub("X", "", year))

pop_ssp1 <- read.csv(file = paste0(dir,"population_ssp1_national_annual_2015_2100.csv")) %>% 
  filter(country == "Australia") %>% 
  select(-iso, -country) %>% 
  gather(value = pop, key = year) %>% 
  mutate(year = gsub("X", "", year)) %>% 
  full_join(pop_hist) %>% 
  mutate(pop = pop*1000000) %>% 
  arrange(year)

pop_ssp5 <- read.csv(file = paste0(dir,"population_ssp5_national_annual_2015_2100.csv")) %>% 
  filter(country == "Australia") %>% 
  select(-iso, -country) %>% 
  gather(value = pop, key = year) %>% 
  mutate(year = gsub("X", "", year)) %>% 
  full_join(pop_hist) %>% 
  mutate(pop = pop*1000000) %>% 
  arrange(year)

# GDP 
# changed format of files: delete first rows for readability: 
#																																																								
# national GDP in PPP 2005 Int $													
# (commas in the country name have been replaced by ; )									
#																

gdp_hist <- read.csv(file = paste0(dir,"gdp-countrylevel_histsoc_annual_1960_2014.csv")) %>% 
  filter(country == "Australia") %>% 
  select(-iso, -country) %>% 
  gather(value = gdp, key = year) %>% 
  mutate(year = gsub("X", "", year))

gdp_ssp1 <- read.csv(file = paste0(dir,"gdp-countrylevel_ssp1_annual_2015_2100.csv")) %>% 
  filter(country == "Australia") %>% 
  select(-iso, -country) %>% 
  gather(value = gdp, key = year) %>% 
  mutate(year = gsub("X", "", year)) %>% 
  full_join(gdp_hist) %>% 
  arrange(year)

gdp_ssp5 <- read.csv(file = paste0(dir,"gdp-countrylevel_ssp5_annual_2015_2100.csv")) %>% 
  filter(country == "Australia") %>% 
  select(-iso, -country) %>% 
  gather(value = gdp, key = year) %>% 
  mutate(year = gsub("X", "", year)) %>% 
  full_join(gdp_hist) %>% 
  arrange(year)

proj_ssp1<-pop_ssp1 %>% 
  full_join(gdp_ssp1) %>% 
  full_join(all_data) %>% 
  mutate(food_extr_2010 = NA, 
         effort_extr_2010 = NA, 
         food_extr_1961 = NA, 
         effort_extr_1961 = NA) %>% 
  filter(year>1960)

proj_ssp5<-pop_ssp5 %>% 
  full_join(gdp_ssp5) %>% 
  full_join(all_data) %>% 
  mutate(food_extr_2010 = NA, 
         effort_extr_2010 = NA, 
         food_extr_1961 = NA, 
         effort_extr_1961 = NA) %>% 
  filter(year>1960)

```

## function to calculate future demand and effort 

```{r}

proj_demand_effort<-function(data){
  
  # data = proj_ssp1
  
  # from Naylor et al 2021
  # # e.g. demand for year 2011 based on 2010
  # # C1 = to estimate 
  # C0 = data %>% filter(year == 2010) %>% select(food)
  # Y0 = data %>% filter(year == 2010) %>% select(pop)
  # Y1 = data %>% filter(year == 2011) %>% select(pop)
  # eps = 0.4
  # G0 = data %>% filter(year == 2010) %>% select(gdp)
  # G1 = data %>% filter(year == 2011) %>% select(gdp)
  # 
  # C1 = C0*(Y1/Y0) + eps*((G1 - G0)/G0)

  eps = 0.2 # here used for calibration
  
  data[1,'food_extr_1961']<-data[1,'food']

  for(i in 1:nrow(data)){
    data[i+1, 'food_extr_1961']<-data[i, 'food_extr_1961']*(data[i+1, 'pop']/data[i, 'pop']) + eps*((data[i+1, 'gdp']-data[i, 'gdp'])/data[i, 'gdp'])
  }

  # add multiplier # NOTE this is not very important as you are using end values of demand that are not scaled.  
  data<-data %>% 
    mutate(food_extr_1961 = food_extr_1961*0.7) %>% 
    filter(!is.na(year))

  # start projections using 2010 as first value 
  data[50,'food_extr_2010']<-data[50,'food']

  for(i in 50:nrow(data)){
    data[i+1, 'food_extr_2010']<-data[i, 'food_extr_2010']*(data[i+1, 'pop']/data[i, 'pop']) + eps*((data[i+1, 'gdp']-data[i, 'gdp'])/data[i, 'gdp'])
  }

  #  multiplier not necessary
  data<-data %>% 
    filter(!is.na(year))

  # final demand trend = FAO data 1961-2010, extrapolation using 2010 as first year 2010-2100
  data<-data %>% 
    mutate(food_final = ifelse(is.na(food), food_extr_2010, food), 
           food_final = food_final*pop) # demand in kg/year

  # calculate effort based on demand and following Naylor's approach 
  # C1 = C0*(Y1/Y0)

  data_eff<-data
  data_eff[1,'effort_extr_1961']<-data_eff[1,'effort']

  for(i in 1:nrow(data_eff)){
    data_eff[i+1, 'effort_extr_1961']<-data_eff[i, 'effort_extr_1961']*(data_eff[i+1, 'food_final']/data_eff[i, 'food_final'])
  }

  # add multiplier - NOTE: this multiplier is important as effort_final is effort 1961-2010 and effort_extr_1961 2010-2100 because we want to make sure historical effort is linked to historical demand.   
  data_eff<-data_eff %>% 
    mutate(effort_extr_1961 = effort_extr_1961*1.5) %>% # NOTE - previously done with food_final instead of food_final*pop, so multiplier was 2.5
    filter(!is.na(year))

  # do the same and calculate from 2010 
  data_eff[50,'effort_extr_2010']<-data_eff[50,'effort']

  for(i in 50:nrow(data_eff)){
    data_eff[i+1, 'effort_extr_2010']<-data_eff[i, 'effort_extr_2010']*(data_eff[i+1, 'food_final']/data_eff[i, 'food_final'])
  }

  data_eff<-data_eff %>% 
    filter(!is.na(year))

  # final effort trend = Effort data 1961-2010, extrapolation using 1961 as first year 2010-2100
  data_eff<-data_eff %>% 
    mutate(effort_final = ifelse(is.na(effort), food_extr_1961, effort))
  
  return(data_eff = data_eff)
  
}

```

## apply function to 2 projections 

This is to apply the function above to both SSPs. The result is demand and effort projections to 2100 under SSP1 and SSP5. 

```{r}

ssp1<-proj_demand_effort(proj_ssp1) %>% 
  mutate(scenarios = ifelse(year <=2010, "Historical", "ssp1"))
ssp5<-proj_demand_effort(proj_ssp5)%>% 
  mutate(scenarios = ifelse(year <=2010, "Historical", "ssp5"))

ssp_all<-ssp1 %>% 
  full_join(ssp5)

```

## plot projections

```{r}

df_plot<-ssp_all %>% 
  gather(value = value, key = data, -year, -scenarios) %>% 
  mutate(origin = case_when(data == "pop" ~ "ISIMIP", 
                            data == "gdp" ~ "ISIMIP", 
                            data == "effort" ~ "Rousseau_et_al_2023", 
                            data == "Food" ~ "FAO", 
                            data == "food_extr_2010" ~ "Extrapolated", 
                            data == "food_extr_1961" ~ "Extrapolated", 
                            data == "effort_extr_2010" ~ "Extrapolated",
                            data == "effort_extr_1961" ~ "Extrapolated",
                            data == "food_final" & year <=2010 ~ "FAO", 
                            data == "food_final" & year >2010 ~ "Extrapolated",
                            data == "effort_final" & year <=2010 ~ "Rousseau_et_al_2023", 
                            data == "effort_final" & year >2010 ~ "Extrapolated"), 
         year = as.numeric(year))
  
# fix scenarios for effort 
df_plot$scenarios<-ifelse(df_plot$data == "effort_extr_1961" & df_plot$year <=2010, "Historical extrapolated", df_plot$scenarios)

# we use food_final which considers the original data from 1961-2010 (food) and the projections from 2011 to 2100 (food_extr_2010) multiplied by population 

plot_demand<-ggplot(df_plot, aes(x = year, y = value))+
  geom_point(data = filter(df_plot, data %in% c("food_final"), origin == "FAO"), aes(color = scenarios), size = 1)+
  geom_line(data = filter(df_plot, data %in% c("food_final"), origin == "Extrapolated"), aes(color = scenarios), size = 1)+
  scale_y_continuous(labels = unit_format(unit = "M", scale = 1e-6))+
  scale_color_manual(name = "Scenarios", values = c("black", "blue", "red"), labels = c("Historical data", "SSP1-mitigation", "SSP5-high-emission"), )+
  my_theme +
  ylab("Demand (kg/y)")+
  xlab("Year")
 
# we use effort_extr_1961 which considers projections between 1961 and 2100 
# we use effort which considers observations between 1961 and 2010 

plot_effort<-ggplot(filter(df_plot, data %in% c("effort", "effort_extr_1961")), aes(x = year, y = value))+
  geom_point(data = filter(df_plot, data %in% c("effort")), aes(color = scenarios), size = 1)+
  geom_line(data = filter(df_plot, data %in% c("effort_extr_1961")), aes(color = scenarios), size = 1)+
  scale_y_continuous(labels = unit_format(unit = "M", scale = 1e-6))+
  scale_color_manual(name = "Scenarios", values = c("black", "grey","blue", "red"), labels = c("Historical data", "Historical projections", "SSP1-mitigation", "SSP5-high-emission"), )+
  my_theme +
  ylab("Effort (DkW)")+
  xlab("Year") 

# remove legend from other plots 
plot_demand<-plot_demand + theme(legend.position = "none")

jpeg("projections_demand_effort.jpg", width = 6, height = 4.5, units = "in", res = 300, bg = "transparent")
plot_demand/plot_effort + plot_layout(guides = "collect") + plot_annotation(tag_levels = 'A')
dev.off()

```

## save data ---- 

```{r}

getwd()
library(data.table)
fwrite(x = df_plot, file.path(dir, "Demand_effort_ssp1_ssp5.csv"))
```


##Creating Effort matrix
```{r}

#ssp1 
Fishing_ssp1<- matrix(NA, nrow = length(unique(ssp1$year)), 
                 ncol = length(unique(Gear_params$gear)),
                 dimnames = list("time" = unique(ssp1$year), 
                                 "effort" = unique(Gear_params$gear)))
  
Fishing_ssp1[,2]<- ssp1$effort_final         
Fishing_ssp1[,1]<- 1

#ssp5

Fishing_ssp5<- matrix(NA, nrow = length(unique(ssp5$year)), 
                 ncol = length(unique(Gear_params$gear)),
                 dimnames = list("time" = unique(ssp5$year), 
                                 "effort" = unique(Gear_params$gear)))
  
Fishing_ssp5[,2]<- ssp5$effort_final         
Fishing_ssp5[,1]<- 1
```

##Forcing the above two effort scenarios on regional model - SouthEast Australia model

```{r}

params2<- newMultispeciesParams( species_params = Species_params, gear_params = Gear_params, kappa = 0.002191539)

sim<- project(params2,t_max=200)
plotBiomass(sim)

params3<- params2

params3@initial_n <- sim@n[200,,]

sim2<- project(params3)
plotBiomass(sim2)

#SSP1
sim_ssp1<- project(params3, effort = Fishing_ssp1)
plot(sim_ssp1)
plotSpectra(sim_ssp1)
plotBiomass(sim_ssp1)

#SSP5
sim_ssp5 <- project(params3, effort = Fishing_ssp5)
plot(sim_ssp5)
plotSpectra(sim_ssp5)
plotBiomass(sim_ssp5)
```



Using R_max to calibrate the model following - North Sea Model Example - mizerHowTo

```{r}
params2<- params
species_params(params2)$erepro <- .01
sim_uncalibrated <- project(params2)
plotCalibration(sim_uncalibrated, stage = 1)
params_guessed <- params2
species_params(params_guessed)$R_max <- resource_params(params_guessed)$kappa *species_params(params_guessed)$w_inf^-1
sim_guessed <- project(params_guessed)
plotCalibration(sim_guessed)

```

Further refining the model 
```{r}
params_optim <- params
vary <-  log10(params_optim@species_params$R_max) 
getError(vary = vary, params = params_optim, dat = params_optim@species_params$biomass_observed)
params_optim<-setParams(params_optim)
noCores <- parallel::detectCores() - 1 
cl <- parallel::makeCluster(noCores, setup_timeout = 0.5)
setDefaultCluster(cl = cl)
clusterExport(cl, varlist = "cl",envir=environment())

clusterEvalQ(cl, {
 library(mizerExperimental)
 library(optimParallel)
  })

optim_result <- optimParallel::optimParallel(par=vary,getError, params=params_optim, dat = params_optim@species_params$biomass_observed, data_type = "SSB", method   ="L-BFGS-B", lower=c(rep(3,dim(params_optim@species_params)[1])), upper= c(rep(15,dim(params_optim@species_params)[1])),
                                             parallel=list(loginfo=TRUE, forward=TRUE))

stopCluster(cl)

params_optim@species_params$R_max <- 10^optim_result$par 
sim_optim <- project(params_optim)

plotSpectra(sim_optim)

```

Functions from Romain's Toothfish model 

```{r}
fastOptim <- function(params)
{
# create set of params for the optimisation process
params_optim <- params
vary <-  log10(params_optim@species_params$R_max) # variable to explore
params_optim<-setParams(params_optim)
# set up workers
noCores <- parallel::detectCores() - 1 # keep some spare core
cl <- parallel::makeCluster(noCores, setup_timeout = 0.5)
setDefaultCluster(cl = cl)
clusterExport(cl, varlist = "cl",envir=environment())
clusterEvalQ(cl, {
  library(mizerExperimental)
  library(optimParallel)
})
optim_result <- optimParallel::optimParallel(par=vary,getError,params=params_optim, dat = params_optim@species_params$biomass_observed, data_type = "SSB", method   ="L-BFGS-B", lower=c(rep(3,dim(params_optim@species_params)[1])), upper= c(rep(15,dim(params_optim@species_params)[1])),
                                             parallel=list(loginfo=TRUE, forward=TRUE))
stopCluster(cl)
params_optim@species_params$R_max <- 10^optim_result$par 
sim_optim <- project(params_optim, t_max = 2000)
return(sim_optim)
}
```


```{r}
# function running tuneParams function in a row for a quick start to a calibration
fastCalib <- function(params, match = F)
{
params <- calibrateBiomass(params) # changes kappa and rmax
if(match) params <- matchBiomasses(params) # set rmax to inf and adjust erepro
params <- steady(params, tol = 0.001)
sim <- project(params, t_max = 1000)
return(sim)
}
```


```{r}
# removing effort = 1 so always using intial effort and removing /1e6 so everything is in grams
getError2 <- function (vary, params, dat, data_type = "catch", tol = 0.1, 
    timetorun = 10) 
{
    params@species_params$R_max[] <- 10^vary[1:length(params@species_params$R_max)]
    params <- setParams(params)
    params <- projectToSteady(params, distance_func = distanceSSLogN, 
        tol = tol, t_max = 200, return_sim = F)
    sim <- project(params, t_max = timetorun, progress_bar = F)
    if (data_type == "SSB") {
        output <- getSSB(sim)[timetorun, ]
    }
    if (data_type == "catch") {
        output <- getYield(sim)[timetorun, ]
    }
    pred <- log(output)
    dat <- log(dat)
    discrep <- pred - dat
    discrep <- (sum(discrep^2))
    return(discrep)
}
```


```{r}
getErrorTime <- function(vary,params,dat,effort,env=state,tol = 0.1) {
  
  params@species_params$R_max[1:9]<-10^vary[1:9]
  params@species_params$erepro[1:9]<-vary[10:18]
  params@gear_params$catchability[1]<-vary[19]
  
  params <- setParams(params)
  # run to steady state and update params
  # env$params<- projectToSteady(env$params, distance_func = distanceSSLogN,
  #                 tol = tol, t_max = 200,return_sim = F)
  
  params_steady<- projectToSteady(params, distance_func = distanceSSLogN,
                                  tol = tol, t_max = 200,return_sim = F)
  
  #run time-varying effort model tthough time with new erepro
  
  simt <- project(params_steady, effort = effort, initial_n =  params_steady@initial_n, initial_n_pp = params_steady@initial_n_pp)
  
  # get biomass through time
  biomass <- sweep(simt@n, 3, simt@params@w * simt@params@dw, "*")
  
  #get yield through time from model:
  
  f_gear<-getFMortGear(params,effort)
  yield_species_gear <- apply(sweep(f_gear, c(1, 3, 4), biomass, "*"),
                              c(1, 2, 3), sum)
  # yield_species_gear
  
  yield_species <-apply(yield_species_gear, c(1, 3), sum)
  
  yield_frame <- melt(yield_species)
  
  # leave out spin up and change units to tonnes    
  # y<-yield_frame[yield_frame$time >= 1947,]
  
  # disregard zeroes - these were NAs only filled in to run the model   
  
  obs<-dat$catchT
  pred<-yield_frame$value[1:18] # only selecting D.ele for now
  
  # sum of squared errors, could use  log-scale of predictions and data (could change this or use other error or likelihood options)
  
  error <- sum((log(pred) - log(obs))^2,na.rm=T)
  
  # can use a strong penalty on the error to ensure we reach a minimum of 10% of the data (biomass or catch) for each species
  # if(any(pred < 0.1*dat)) discrep <- discrep + 1e10
  
  return(error)
  
}
```


### Set of functions using size selectivity of the data

```{r}
# this function adds a lower boundary to selected size
plotBiomassObservedVsModelCustom <- function (object, species = NULL, ratio = FALSE, log_scale = TRUE, 
    return_data = FALSE, labels = TRUE, show_unobserved = FALSE) 
{
    if (is(object, "MizerSim")) {
        params = object@params
        n <- finalN(object)
    }
    else if (is(object, "MizerParams")) {
        params = object
        n <- initialN(params)
    }
    else {
        stop("You have not provided a valid mizerSim or mizerParams object.")
    }
    sp_params <- params@species_params
    species = valid_species_arg(object, species)
    if (length(species) == 0) 
        stop("No species selected, please fix.")
    row_select = match(species, sp_params$species)
    if (!"biomass_observed" %in% names(sp_params)) {
        stop("You have not provided values for the column 'biomass_observed' ", 
            "in the mizerParams/mizerSim object.")
    }
    else if (!is.numeric(sp_params$biomass_observed)) {
        stop("The column 'biomass_observed' in the mizerParams/mizerSim object", 
            " is not numeric, please fix.")
    }
    else {
        biomass_observed = sp_params$biomass_observed
    }
    
    cutoffLow <- sp_params$biomass_cutoffLow[row_select]
    if (is.null(cutoffLow)) {
        cutoffLow = rep(0, length(species))
    }
    else if (!is.numeric(cutoffLow)) {
        stop("params@species_params$biomass_cutoffLow is not numeric, \",\n                 \"please fix.")
    }
    cutoffLow[is.na(cutoffLow)] <- 0
    
    cutoffHigh <- sp_params$biomass_cutoffHigh[row_select]
    if (is.null(cutoffHigh)) {
        cutoffHigh = rep(0, length(species))
    }
    else if (!is.numeric(cutoffHigh)) {
        stop("params@species_params$biomass_cutoffHigh is not numeric, \",\n                 \"please fix.")
    }
    cutoffHigh[is.na(cutoffHigh)] <- 0
    
    sim_biomass = rep(0, length(species))
    for (j in 1:length(species)) {
        sim_biomass[j] = sum((n[row_select[j], ] * params@w * 
            params@dw)[params@w >= cutoffLow[j] & cutoffHigh[j] >= params@w])
    }
    dummy = data.frame(species = species, model = sim_biomass, 
        observed = biomass_observed[row_select]) %>% mutate(species = factor(species, 
        levels = species), is_observed = !is.na(observed) & observed > 
        0, observed = case_when(is_observed ~ observed, !is_observed ~ 
        model), ratio = model/observed)
    if (sum(dummy$is_observed) == 0) {
        cat(paste("There are no observed biomasses to compare to model,", 
            "only plotting model biomasses.", sep = "\n"))
    }
    if (!show_unobserved) {
        dummy <- filter(dummy, is_observed)
    }
    if (return_data == TRUE) 
        return(dummy)
    tre <- round(sum(abs(1 - dummy$ratio)), digits = 3)
    caption <- paste0("Total relative error = ", tre)
    if (any(!dummy$is_observed)) {
        caption <- paste(caption, "\n Open circles represent species without biomass observation.")
    }
    if (ratio == FALSE) {
        gg <- ggplot(data = dummy, aes(x = observed, y = model, 
            colour = species, shape = is_observed)) + geom_abline(aes(intercept = 0, 
            slope = 1), colour = "purple", linetype = "dashed", 
            size = 1.3) + geom_point(size = 3) + labs(y = "model biomass [g]") + 
            coord_cartesian(ylim = range(dummy$model, dummy$observed))
    }
    else {
        gg <- ggplot(data = dummy, aes(x = observed, y = ratio, 
            colour = species, shape = is_observed)) + geom_hline(aes(yintercept = 1), 
            linetype = "dashed", colour = "purple", 
            size = 1.3) + geom_point(size = 3) + labs(y = "model biomass / observed biomass") + 
            coord_cartesian(ylim = range(dummy$ratio))
    }
    gg <- gg + labs(x = "observed biomass [g]", caption = caption) + 
        scale_colour_manual(values = getColours(params)[dummy$species]) + 
        scale_shape_manual(values = c(`TRUE` = 19, `FALSE` = 1)) + 
        guides(shape = "none")
    if (log_scale == TRUE & ratio == FALSE) {
        gg = gg + scale_x_log10() + scale_y_log10()
    }
    if (log_scale == TRUE & ratio == TRUE) {
        gg = gg + scale_x_log10()
    }
    if (labels == TRUE) {
        gg = gg + ggrepel::geom_label_repel(aes(label = species), 
            box.padding = 0.35, point.padding = 0.5, segment.color = "grey50", 
            show.legend = FALSE, max.overlaps = Inf, seed = 42)
    }
    gg
}
```


```{r}
# adapting cutoff here too
calibrateBiomassCustom <- function (params) 
{
  if ((!("biomass_observed" %in% names(params@species_params))) || 
      all(is.na(params@species_params$biomass_observed))) {
    return(params)
  }
  no_sp <- nrow(params@species_params)
  
  cutoffLow <- params@species_params$biomass_cutoffLow
  if (is.null(cutoffLow)) 
    cutoffLow <- rep(0, no_sp)
  cutoffLow[is.na(cutoffLow)] <- 0
  
  cutoffHigh <- params@species_params$biomass_cutoffHigh
  if (is.null(cutoffHigh)) 
    cutoffHigh <- rep(0, no_sp)
  cutoffHigh[is.na(cutoffHigh)] <- 0
  
  observed <- params@species_params$biomass_observed
  observed_total <- sum(observed, na.rm = TRUE)
  sp_observed <- which(!is.na(observed))
  model_total <- 0
  for (sp_idx in sp_observed) {
    model_total <- model_total + sum((params@initial_n[sp_idx, 
    ] * params@w * params@dw)[params@w >= cutoffLow[sp_idx] & cutoffHigh[sp_idx] >= params@w])
  }
  scaleModel(params, factor = observed_total/model_total)
}
```

```{r}
# same as above
matchBiomassCustom <- function (params, species = NULL) 
{
    if (!("biomass_observed" %in% names(params@species_params))) {
        return(params)
    }
    species <- valid_species_arg(params, species = species, return.logical = TRUE) & 
        !is.na(params@species_params$biomass_observed) & params@species_params$biomass_observed > 
        0
    for (sp in (1:nrow(params@species_params))[species]) {
        cutoffLow <- params@species_params$biomass_cutoffLow[[sp]]
        if (is.null(cutoffLow) || is.na(cutoffLow)) {
            cutoffLow <- 0
        }
                    cutoffHigh <- params@species_params$biomass_cutoffHigh[[sp]]
        if (is.null(cutoffHigh) || is.na(cutoffHigh)) {
            cutoffHigh <- 0
        }
        
        total <- sum((params@initial_n[sp, ] * params@w * params@dw)[params@w >= cutoffLow & cutoffHigh >= params@w])
        factor <- params@species_params$biomass_observed[[sp]]/total
        params@initial_n[sp, ] <- params@initial_n[sp, ] * factor
    }
    setBevertonHolt(params)
}
```


```{r}
getErrorCustom <- function(vary, params, dat, tol = 0.001, 
    timetorun = 10)
{
  params@species_params$R_max[1:9]<-10^vary[1:9]
  params@species_params$erepro[1:9]<-vary[10:18]
  params@species_params$interaction_resource[1:9] <- vary[19:27]
  params <- setParams(params)
  interaction <- params@interaction
  interaction[] <- matrix(vary[28:108],nrow = 9) # stop at 54 if looking only at 3 biggest species
  
  params <- setInteraction(params,interaction)
    params <- projectToSteady(params, distance_func = distanceSSLogN, 
        tol = tol, t_max = 200, return_sim = F)
    
    sim <- project(params, t_max = timetorun, progress_bar = F)
    
    sim_biomass = rep(0, length(params@species_params$species))
    
        cutoffLow <- params@species_params$biomass_cutoffLow
    if (is.null(cutoffLow)) 
        cutoffLow <- rep(0, no_sp)
    cutoffLow[is.na(cutoffLow)] <- 0
    
        cutoffHigh <- params@species_params$biomass_cutoffHigh
    if (is.null(cutoffHigh)) 
        cutoffHigh <- rep(0, no_sp)
    cutoffHigh[is.na(cutoffHigh)] <- 0
        
    for (j in 1:length(sim_biomass)) {
        sim_biomass[j] = sum((sim@n[dim(sim@n)[1],j,] * params@w * 
            params@dw)[params@w >= cutoffLow[j] & cutoffHigh[j] >= params@w])
    }
    
    pred <- log(sim_biomass)
    dat <- log(dat)
    discrep <- pred - dat
    discrep <- (sum(discrep^2))
    return(discrep)
}
```
```

```{r}
Param=DFparam 
effort_histsoc_1950_2014<- read_csv("effort_histsoc_1950_2014.csv")


```

#Data 

