---
title: SEA Model 
author: NR
editor_options: 
  chunk_output_type: console
---

Setting up Libraries

```{r}
#Load data 
rm(list=ls())

#Load allll the libraries - use install packages if library doesn't work
library(readxl)
library(tidyverse)
library(devtools)
library(plyr) 
library(Rcpp)
library(reshape2) 
library(inline)
library(ggplot2)
library(ggthemes)
library(mizer)
library(ggpubr)
library(mizerMR)
library(mizerExperimental)
library(readxl)
library(mizerHowTo)
library(mgcv)
library(scales)
library(patchwork)

summarise <-dplyr::summarise

my_theme<-theme_bw()+
  theme(text = element_text(size = 10), # this should be overwritten by the below
        plot.title = element_text(size = 10),
        axis.title = element_text(size = 9),
        axis.text = element_text(size = 6),
        legend.title=element_text(size = 9),
        legend.text=element_text(size = 8),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.key.size = unit(0.1, "cm"))



dir = "/Users/raveendransankaran/Documents/prepare-Nilar14/"
```

Loading param from Cami and formatting it to Mizer
```{r}
DFparam <- read_excel("mixed fihseries data/DFparam.xlsx")

Species_params <- DFparam[,(-c(1,10,15,16,17,18,19,20,21,25,26,27,28,29,30,32,34,35,36,37))]
Species_params$R_max <- Species_params$r_max


Species_params<- Species_params[,(-c(15))]
Species_params$biomass_observed <- Species_params$ssbObs
avail_pp<- c(1.0, 0.2, 0.8, 0.2, 0.8, 0.2, 0.0, 0.3, 0.2, 0.2, 0.3, 0.2, 0.2, 0.3, 0.2, 0.2, 0.0, 0.0, 0.0)
avail_bb <- c(0.0, 0.4, 0.0, 0.4, 0.0, 0.4, 0.8, 0.5, 0.4, 0.4, 0.5, 0.4, 0.4, 0.5, 0.4, 0.4, 0.8, 0.8, 0.8)
Species_params$interaction_resource<- avail_pp+avail_bb
Species_params$ssbObs <- NULL

Species_params$l_inf <- exp(log(Species_params$w_inf/Species_params$a)/Species_params$b)

Species_params$FGroup<- c("bathypelagic<30cm", "demersal30-90cm", "demersal30-90cm","bathydemersal30-90cm", "benthopelagic30-90cm", "benthopelagic30-90cm", "shark>=90cm","demersal30-90cm","flatfish<90cm", "benthopelagic30-90cm", "benthopelagic30-90cm", "bathypelagic30-90cm", "benthopelagic30-90cm", "benthopelagic>=90cm", "benthopelagic>=90cm", "bathydemersal>=90cm", "shark>=90cm", "shark>=90cm", "shark>=90cm" )

# Fishing params 
Gear_params <- DFparam[,c(2,17,18,19,20,21,28)]

trawl <- function(w,W25_S,W50_S,...){

  SR <- W50_S - W25_S
  S1 <- W50_S*log(3)/SR
  S2 <- S1 / W50_S
  sel <-1 / (1 + exp(S1 - S2*w))
  return(sel)
}

gillnet<-function(w,W50_B,sigma_B,...){
  sel<-exp(-(w-W50_B)^2/sigma_B^2)
  return(sel)
}

# not working for now ....
unique(Gear_params$species)
demersal <-c("centroberyx affinis",  "nemadactylus macropterus","platycephalus richardsoni", "zeus faber" , "seriolella brama","hoplostethus atlanticus", "macruronus novaezelandiae", "seriolella punctata", "rexea solandri","genypterus blacodes")
pelagic <- c("myctophids","sillago flindersi","nototodarus gouldi","helicolenus barathri","trachurus declivis","squalus spp.", "pristiophorus cirratus","mustelus antarcticus", "galeorhinus galeus" )

str(Gear_params)
Gear_params$gear=ifelse(Gear_params$species %in% demersal, "demersal", "pelagic")
# end not working 
 
Gear_params$gear<-"pelagic"
Gear_params[c(6,8:16),'gear']<-"demersal"
head(Gear_params)
```



```{r}
#Gear_params$sel_func<- "sigmoid_length" 

#Gear_params$l25<- exp(log(Gear_params$W25_S/Species_params$a)/Species_params$b)
#Gear_params$l50<- exp(log(Gear_params$W50_S/Species_params$a)/Species_params$b)



params <- newMultispeciesParams(species_params = Species_params, gear_params = Gear_params, kappa = 0.002191539)


```

```{r}
plotSpectra(params)
Simparam <- project(params, effort = 1)
plotSpectra(Simparam)
plotBiomass(Simparam)


```

Code here shows that the biomass is decreasing drastically - need to change the R_max value
```{r}

params@species_params$R_max<- params@resource_params$kappa*params@species_params$w_inf^-1
Simparam<-project(params, effort = 1)
plotSpectra(Simparam)
plotBiomass(Simparam)

```

All Species Co-exist : Except one species - Squalus spp.

```{r}
plotFeedingLevel(Simparam, include_critical = TRUE)
plotGrowthCurves(Simparam, species_panel = TRUE)
plotlySpectra(Simparam)

plotlyPredMort(Simparam)
plotDiet(Simparam)#, species = "squalus spp.")
plotEnergyBudget(Simparam)
```

Loading and plotting : catch and effort (Fish-MIP data)


```{r}
effort_his <- read.csv(file = "effort_histsoc_1841_2010_regional_models.csv") 
catch_his <- read.csv(file = "calibration_catch_histsoc_1850_2004_regional_models.csv")

effort_SEA <- filter(effort_his,region=="SEAustralia.mizer")#subset - only industrial , all fishing countries, gear - trawl bottom and gillnet (check the  Fgroups), plot per gear and Fgroups- facet_wrap().

effort_tot <- effort_SEA


effort_tot<- effort_tot %>% 
  group_by(Year,FGroup) %>% 
  dplyr::summarise(TotalEffort=sum(NomActive)) %>% 
  ungroup()

 head(effort_tot)

ggplot(data = effort_tot, aes(x=Year,y=TotalEffort, color = FGroup))+
  geom_line()

catch_SEA <- filter(catch_his, region=="SEAustralia.mizer")

catch_tot<- catch_SEA
catch_tot$Totalcatches <- catch_tot$Reported+catch_tot$IUU+catch_tot$Discards
 head(catch_tot)#checked-works
 
catch_tot <- catch_tot %>%
  group_by(Year) %>%
  dplyr::summarise(Totalcatches=sum(Totalcatches))

ggplot(data = catch_tot, aes(x=Year,y=Totalcatches))+
  geom_line()

#ggplot(data = effort_SEA, aes(x= Year, y= NomActive, group = 1))
#+geom_line(linetype="solid", color="forestgreen")+ theme_minimal()+labs(x="Year" ,y="Fishing effort", title = "Fishing effort (1841-2010)- SE-A")+theme(plot.title = element_text(hjust = 0.5,size = 10,face = "bold"))

#ggplot(data = catch_SEA, aes(x= Year, y= Reported, group = 1))+geom_line(linetype="solid", color="turquoise4")+ theme_minimal()+labs(x="Year" ,y="Catch (reported)", title = "Catches(1920-2004)- SE-A")+theme(plot.title = element_text(hjust = 0.5,size = 10,face = "bold"))



```
Creating a subset of effort and catch from the above fishmip data - with only industrial sector and trawl bottom and gillnet gear.

```{r}

# try filter only cephalopods, demersal, bathydemersal, flatfish, rays, shark

effort_sub <- filter(effort_SEA, Sector == "Industrial",Gear =="Trawl_Bottom" | Gear == "Gillnets", FGroup %in% c("cephalopods", "bathydemersal<30cm", "bathydemersal>=90cm","bathydemersal30-90cm", "demersal<30cm", "demersal>=90cm", "demersal30-90cm", "flatfish<90cm", "rays<90cm", "rays>=90cm", "shark>=90cm", "bathypelagic<30cm", "bathypelagic30-90cm", "benthopelagic30-90cm", "benthopelagic>=90cm"))
# Gear %in% c("Trawl_Bottom", "Gillnets")

head(effort_sub)

effort_tt<- effort_sub


effort_tt<- effort_tt %>% 
  group_by(Year,Gear,FGroup) %>% 
  dplyr::summarise(TotalEffort=sum(NomActive)) %>%
  ungroup()

head(effort_tt)

ggplot(data = effort_tt, aes(x = Year, y = TotalEffort,color = FGroup)) +
  geom_line( size = 1) +
  labs(title = "Gear") + 
  facet_wrap(~ Gear)


catch_sub <- filter(catch_SEA,FGroup %in% c("cephalopods", "bathydemersal<30cm", "bathydemersal>=90cm","bathydemersal30-90cm", "demersal<30cm", "demersal>=90cm", "demersal30-90cm", "flatfish<90cm", "rays<90cm", "rays>=90cm", "shark>=90cm", "bathypelagic<30cm", "bathypelagic30-90cm", "benthopelagic30-90cm", "benthopelagic>=90cm"))

head(catch_sub)

catch_tt<- catch_sub
catch_tt$Totalcatches <- catch_tt$Reported+catch_tt$IUU+catch_tt$Discards

catch_tt<- catch_tt %>%
  group_by(Year,FGroup) %>%
  dplyr::summarise(Totalcatches=sum(Totalcatches)) %>%
  ungroup()

head(catch_tt)

ggplot(data = catch_tt, aes(x = Year, y = Totalcatches,color = FGroup)) +
  geom_line( size = 1) +
  labs(title = "FGroup") 
  
#ggplot(data = catch_tt, aes(x = Year, y = Totalcatches,color = FGroup)) + geom_line( size = 1) +labs(title = "SAUP") +facet_wrap(~ SAUP) - each SAUP has only 1 observation - too many plots with too little data .  


  #  platycephalus - flatfish , 17-19,7 - sharks, 3 - ceph, 6- demersal <30, - match species to Fgroup and gear. - set up gear params- gear/year- 18
 

```


##Creating Effort matrix
```{r}

#Using effort calculated using demand from script developed by Camilla Novaglio 
#ssp1 
Fishing_ssp1<- matrix(NA, nrow = length(unique(ssp1$year)), 
                 ncol = length(unique(Gear_params$gear)),
                 dimnames = list("time" = unique(ssp1$year), 
                                 "effort" = unique(Gear_params$gear)))

  
Fishing_ssp1[,2]<- ssp1$effort_final         
Fishing_ssp1[,1]<- ssp1_p$effort_final

#ssp5

Fishing_ssp5<- matrix(NA, nrow = length(unique(ssp5$year)), 
                 ncol = length(unique(Gear_params$gear)),
                 dimnames = list("time" = unique(ssp5$year), 
                                 "effort" = unique(Gear_params$gear)))
  
Fishing_ssp5[,2]<- ssp5$effort_final         
Fishing_ssp5[,1]<- ssp5_p$effort_final
```

##Forcing the above two effort scenarios on regional model - SouthEast Australia model

```{r}

params2<- newMultispeciesParams( species_params = Species_params, gear_params = Gear_params, kappa = 0.002191539)

sim<- project(params2,t_max=200)
plotBiomass(sim)

params3<- params2

params3@initial_n <- sim@n[200,,]

sim2<- project(params3)
plotBiomass(sim2)

#SSP1
sim_ssp1<- project(params3, effort = Fishing_ssp1)
plot(sim_ssp1)
plotSpectra(sim_ssp1)
plotBiomass(sim_ssp1)

#SSP5
sim_ssp5 <- project(params3, effort = Fishing_ssp5)
plot(sim_ssp5)
plotSpectra(sim_ssp5)
plotBiomass(sim_ssp5)


# convert the unit of effort:
# 1 divide by the volume of SEA region to get day at seaxkwxm3

Fishing_ssp1_scaled <-Fishing_ssp1
Fishing_ssp1_scaled[,1]<-Fishing_ssp1_scaled[,1]*10^-6
Fishing_ssp1_scaled[,2]<-Fishing_ssp1_scaled[,2]*10^-6

# 2 use catchability to scale down effort and get mortality value 
params4<-params3
params4@gear_params[2,7]<-params3@gear_params[2,7]*0.1
params4@gear_params[3,7]<-params3@gear_params[3,7]*0.00001

# remember to consider only forcings to 2004 (the first 44 years of effort)
#SSP1
sim_ssp1_scaled<- project(params4, effort = Fishing_ssp1_scaled[1:44,])
# plot(sim_ssp1_scaled)
plotBiomass(sim_ssp1_scaled)
plotFMort(sim_ssp1_scaled)
plotYield(sim_ssp1_scaled)

# 3 once you have coexistence, compare modelled catches to empirical catches using the function getErrorCustom() in MizerHow - ASK ROMAIN at this point.  
```


Using R_max to calibrate the model following - North Sea Model Example - mizerHowTo

```{r}
params2<- params
species_params(params2)$erepro <- .01
sim_uncalibrated <- project(params2)
plotCalibration(sim_uncalibrated, stage = 1)
params_guessed <- params2
species_params(params_guessed)$R_max <- resource_params(params_guessed)$kappa *species_params(params_guessed)$w_inf^-1
sim_guessed <- project(params_guessed)
plotCalibration(sim_guessed)

```

Further refining the model 
```{r}
params_optim <- params
vary <-  log10(params_optim@species_params$R_max) 
getError(vary = vary, params = params_optim, dat = params_optim@species_params$biomass_observed)
params_optim<-setParams(params_optim)
noCores <- parallel::detectCores() - 1 
cl <- parallel::makeCluster(noCores, setup_timeout = 0.5)
setDefaultCluster(cl = cl)
clusterExport(cl, varlist = "cl",envir=environment())

clusterEvalQ(cl, {
 library(mizerExperimental)
 library(optimParallel)
  })

optim_result <- optimParallel::optimParallel(par=vary,getError, params=params_optim, dat = params_optim@species_params$biomass_observed, data_type = "SSB", method   ="L-BFGS-B", lower=c(rep(3,dim(params_optim@species_params)[1])), upper= c(rep(15,dim(params_optim@species_params)[1])),
                                             parallel=list(loginfo=TRUE, forward=TRUE))

stopCluster(cl)

params_optim@species_params$R_max <- 10^optim_result$par 
sim_optim <- project(params_optim)

plotSpectra(sim_optim)

```

Functions from Romain's Toothfish model 

```{r}
fastOptim <- function(params)
{
# create set of params for the optimisation process
params_optim <- params
vary <-  log10(params_optim@species_params$R_max) # variable to explore
params_optim<-setParams(params_optim)
# set up workers
noCores <- parallel::detectCores() - 1 # keep some spare core
cl <- parallel::makeCluster(noCores, setup_timeout = 0.5)
setDefaultCluster(cl = cl)
clusterExport(cl, varlist = "cl",envir=environment())
clusterEvalQ(cl, {
  library(mizerExperimental)
  library(optimParallel)
})
optim_result <- optimParallel::optimParallel(par=vary,getError,params=params_optim, dat = params_optim@species_params$biomass_observed, data_type = "SSB", method   ="L-BFGS-B", lower=c(rep(3,dim(params_optim@species_params)[1])), upper= c(rep(15,dim(params_optim@species_params)[1])),
                                             parallel=list(loginfo=TRUE, forward=TRUE))
stopCluster(cl)
params_optim@species_params$R_max <- 10^optim_result$par 
sim_optim <- project(params_optim, t_max = 2000)
return(sim_optim)
}
```


```{r}
# function running tuneParams function in a row for a quick start to a calibration
fastCalib <- function(params, match = F)
{
params <- calibrateBiomass(params) # changes kappa and rmax
if(match) params <- matchBiomasses(params) # set rmax to inf and adjust erepro
params <- steady(params, tol = 0.001)
sim <- project(params, t_max = 1000)
return(sim)
}
```


```{r}
# removing effort = 1 so always using intial effort and removing /1e6 so everything is in grams
getError2 <- function (vary, params, dat, data_type = "catch", tol = 0.1, 
    timetorun = 10) 
{
    params@species_params$R_max[] <- 10^vary[1:length(params@species_params$R_max)]
    params <- setParams(params)
    params <- projectToSteady(params, distance_func = distanceSSLogN, 
        tol = tol, t_max = 200, return_sim = F)
    sim <- project(params, t_max = timetorun, progress_bar = F)
    if (data_type == "SSB") {
        output <- getSSB(sim)[timetorun, ]
    }
    if (data_type == "catch") {
        output <- getYield(sim)[timetorun, ]
    }
    pred <- log(output)
    dat <- log(dat)
    discrep <- pred - dat
    discrep <- (sum(discrep^2))
    return(discrep)
}
```


```{r}
getErrorTime <- function(vary,params,dat,effort,env=state,tol = 0.1) {
  
  params@species_params$R_max[1:9]<-10^vary[1:9]
  params@species_params$erepro[1:9]<-vary[10:18]
  params@gear_params$catchability[1]<-vary[19]
  
  params <- setParams(params)
  # run to steady state and update params
  # env$params<- projectToSteady(env$params, distance_func = distanceSSLogN,
  #                 tol = tol, t_max = 200,return_sim = F)
  
  params_steady<- projectToSteady(params, distance_func = distanceSSLogN,
                                  tol = tol, t_max = 200,return_sim = F)
  
  #run time-varying effort model tthough time with new erepro
  
  simt <- project(params_steady, effort = effort, initial_n =  params_steady@initial_n, initial_n_pp = params_steady@initial_n_pp)
  
  # get biomass through time
  biomass <- sweep(simt@n, 3, simt@params@w * simt@params@dw, "*")
  
  #get yield through time from model:
  
  f_gear<-getFMortGear(params,effort)
  yield_species_gear <- apply(sweep(f_gear, c(1, 3, 4), biomass, "*"),
                              c(1, 2, 3), sum)
  # yield_species_gear
  
  yield_species <-apply(yield_species_gear, c(1, 3), sum)
  
  yield_frame <- melt(yield_species)
  
  # leave out spin up and change units to tonnes    
  # y<-yield_frame[yield_frame$time >= 1947,]
  
  # disregard zeroes - these were NAs only filled in to run the model   
  
  obs<-dat$catchT
  pred<-yield_frame$value[1:18] # only selecting D.ele for now
  
  # sum of squared errors, could use  log-scale of predictions and data (could change this or use other error or likelihood options)
  
  error <- sum((log(pred) - log(obs))^2,na.rm=T)
  
  # can use a strong penalty on the error to ensure we reach a minimum of 10% of the data (biomass or catch) for each species
  # if(any(pred < 0.1*dat)) discrep <- discrep + 1e10
  
  return(error)
  
}
```


### Set of functions using size selectivity of the data

```{r}
# this function adds a lower boundary to selected size
plotBiomassObservedVsModelCustom <- function (object, species = NULL, ratio = FALSE, log_scale = TRUE, 
    return_data = FALSE, labels = TRUE, show_unobserved = FALSE) 
{
    if (is(object, "MizerSim")) {
        params = object@params
        n <- finalN(object)
    }
    else if (is(object, "MizerParams")) {
        params = object
        n <- initialN(params)
    }
    else {
        stop("You have not provided a valid mizerSim or mizerParams object.")
    }
    sp_params <- params@species_params
    species = valid_species_arg(object, species)
    if (length(species) == 0) 
        stop("No species selected, please fix.")
    row_select = match(species, sp_params$species)
    if (!"biomass_observed" %in% names(sp_params)) {
        stop("You have not provided values for the column 'biomass_observed' ", 
            "in the mizerParams/mizerSim object.")
    }
    else if (!is.numeric(sp_params$biomass_observed)) {
        stop("The column 'biomass_observed' in the mizerParams/mizerSim object", 
            " is not numeric, please fix.")
    }
    else {
        biomass_observed = sp_params$biomass_observed
    }
    
    cutoffLow <- sp_params$biomass_cutoffLow[row_select]
    if (is.null(cutoffLow)) {
        cutoffLow = rep(0, length(species))
    }
    else if (!is.numeric(cutoffLow)) {
        stop("params@species_params$biomass_cutoffLow is not numeric, \",\n                 \"please fix.")
    }
    cutoffLow[is.na(cutoffLow)] <- 0
    
    cutoffHigh <- sp_params$biomass_cutoffHigh[row_select]
    if (is.null(cutoffHigh)) {
        cutoffHigh = rep(0, length(species))
    }
    else if (!is.numeric(cutoffHigh)) {
        stop("params@species_params$biomass_cutoffHigh is not numeric, \",\n                 \"please fix.")
    }
    cutoffHigh[is.na(cutoffHigh)] <- 0
    
    sim_biomass = rep(0, length(species))
    for (j in 1:length(species)) {
        sim_biomass[j] = sum((n[row_select[j], ] * params@w * 
            params@dw)[params@w >= cutoffLow[j] & cutoffHigh[j] >= params@w])
    }
    dummy = data.frame(species = species, model = sim_biomass, 
        observed = biomass_observed[row_select]) %>% mutate(species = factor(species, 
        levels = species), is_observed = !is.na(observed) & observed > 
        0, observed = case_when(is_observed ~ observed, !is_observed ~ 
        model), ratio = model/observed)
    if (sum(dummy$is_observed) == 0) {
        cat(paste("There are no observed biomasses to compare to model,", 
            "only plotting model biomasses.", sep = "\n"))
    }
    if (!show_unobserved) {
        dummy <- filter(dummy, is_observed)
    }
    if (return_data == TRUE) 
        return(dummy)
    tre <- round(sum(abs(1 - dummy$ratio)), digits = 3)
    caption <- paste0("Total relative error = ", tre)
    if (any(!dummy$is_observed)) {
        caption <- paste(caption, "\n Open circles represent species without biomass observation.")
    }
    if (ratio == FALSE) {
        gg <- ggplot(data = dummy, aes(x = observed, y = model, 
            colour = species, shape = is_observed)) + geom_abline(aes(intercept = 0, 
            slope = 1), colour = "purple", linetype = "dashed", 
            size = 1.3) + geom_point(size = 3) + labs(y = "model biomass [g]") + 
            coord_cartesian(ylim = range(dummy$model, dummy$observed))
    }
    else {
        gg <- ggplot(data = dummy, aes(x = observed, y = ratio, 
            colour = species, shape = is_observed)) + geom_hline(aes(yintercept = 1), 
            linetype = "dashed", colour = "purple", 
            size = 1.3) + geom_point(size = 3) + labs(y = "model biomass / observed biomass") + 
            coord_cartesian(ylim = range(dummy$ratio))
    }
    gg <- gg + labs(x = "observed biomass [g]", caption = caption) + 
        scale_colour_manual(values = getColours(params)[dummy$species]) + 
        scale_shape_manual(values = c(`TRUE` = 19, `FALSE` = 1)) + 
        guides(shape = "none")
    if (log_scale == TRUE & ratio == FALSE) {
        gg = gg + scale_x_log10() + scale_y_log10()
    }
    if (log_scale == TRUE & ratio == TRUE) {
        gg = gg + scale_x_log10()
    }
    if (labels == TRUE) {
        gg = gg + ggrepel::geom_label_repel(aes(label = species), 
            box.padding = 0.35, point.padding = 0.5, segment.color = "grey50", 
            show.legend = FALSE, max.overlaps = Inf, seed = 42)
    }
    gg
}
```


```{r}
# adapting cutoff here too
calibrateBiomassCustom <- function (params) 
{
  if ((!("biomass_observed" %in% names(params@species_params))) || 
      all(is.na(params@species_params$biomass_observed))) {
    return(params)
  }
  no_sp <- nrow(params@species_params)
  
  cutoffLow <- params@species_params$biomass_cutoffLow
  if (is.null(cutoffLow)) 
    cutoffLow <- rep(0, no_sp)
  cutoffLow[is.na(cutoffLow)] <- 0
  
  cutoffHigh <- params@species_params$biomass_cutoffHigh
  if (is.null(cutoffHigh)) 
    cutoffHigh <- rep(0, no_sp)
  cutoffHigh[is.na(cutoffHigh)] <- 0
  
  observed <- params@species_params$biomass_observed
  observed_total <- sum(observed, na.rm = TRUE)
  sp_observed <- which(!is.na(observed))
  model_total <- 0
  for (sp_idx in sp_observed) {
    model_total <- model_total + sum((params@initial_n[sp_idx, 
    ] * params@w * params@dw)[params@w >= cutoffLow[sp_idx] & cutoffHigh[sp_idx] >= params@w])
  }
  scaleModel(params, factor = observed_total/model_total)
}
```

```{r}
# same as above
matchBiomassCustom <- function (params, species = NULL) 
{
    if (!("biomass_observed" %in% names(params@species_params))) {
        return(params)
    }
    species <- valid_species_arg(params, species = species, return.logical = TRUE) & 
        !is.na(params@species_params$biomass_observed) & params@species_params$biomass_observed > 
        0
    for (sp in (1:nrow(params@species_params))[species]) {
        cutoffLow <- params@species_params$biomass_cutoffLow[[sp]]
        if (is.null(cutoffLow) || is.na(cutoffLow)) {
            cutoffLow <- 0
        }
                    cutoffHigh <- params@species_params$biomass_cutoffHigh[[sp]]
        if (is.null(cutoffHigh) || is.na(cutoffHigh)) {
            cutoffHigh <- 0
        }
        
        total <- sum((params@initial_n[sp, ] * params@w * params@dw)[params@w >= cutoffLow & cutoffHigh >= params@w])
        factor <- params@species_params$biomass_observed[[sp]]/total
        params@initial_n[sp, ] <- params@initial_n[sp, ] * factor
    }
    setBevertonHolt(params)
}
```


```{r}
getErrorCustom <- function(vary, params, dat, tol = 0.001, 
    timetorun = 10)
{
  params@species_params$R_max[1:9]<-10^vary[1:9]
  params@species_params$erepro[1:9]<-vary[10:18]
  params@species_params$interaction_resource[1:9] <- vary[19:27]
  params <- setParams(params)
  interaction <- params@interaction
  interaction[] <- matrix(vary[28:108],nrow = 9) # stop at 54 if looking only at 3 biggest species
  
  params <- setInteraction(params,interaction)
    params <- projectToSteady(params, distance_func = distanceSSLogN, 
        tol = tol, t_max = 200, return_sim = F)
    
    sim <- project(params, t_max = timetorun, progress_bar = F)
    
    sim_biomass = rep(0, length(params@species_params$species))
    
        cutoffLow <- params@species_params$biomass_cutoffLow
    if (is.null(cutoffLow)) 
        cutoffLow <- rep(0, no_sp)
    cutoffLow[is.na(cutoffLow)] <- 0
    
        cutoffHigh <- params@species_params$biomass_cutoffHigh
    if (is.null(cutoffHigh)) 
        cutoffHigh <- rep(0, no_sp)
    cutoffHigh[is.na(cutoffHigh)] <- 0
        
    for (j in 1:length(sim_biomass)) {
        sim_biomass[j] = sum((sim@n[dim(sim@n)[1],j,] * params@w * 
            params@dw)[params@w >= cutoffLow[j] & cutoffHigh[j] >= params@w])
    }
    
    pred <- log(sim_biomass)
    dat <- log(dat)
    discrep <- pred - dat
    discrep <- (sum(discrep^2))
    return(discrep)
}
```
```

```{r}
Param=DFparam 
effort_histsoc_1950_2014<- read_csv("effort_histsoc_1950_2014.csv")


```

#Data 

